---
title: "Project CPT_S"
author: "Khalil Cherif"
date: "2024-10-24"
output: pdf_document
---

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Load the building metadata file
setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")
building_metadata <- read.csv("building_metadata.csv")

# Load the energy consumption file (train or test data)
setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")
energy_data <- read.csv("train 1-2.csv")  # Adjust file path as necessary

# Merge the building metadata and energy data on the 'building_id' column
merged_data <- merge(building_metadata, energy_data, by = "building_id")

# Summarize the meter readings by building_id to get total energy consumption
summarized_data <- merged_data %>%
  group_by(building_id, square_feet) %>%
  summarize(total_energy_consumption = sum(meter_reading, na.rm = TRUE))

# Plot square footage against total energy consumption using ggplot2
ggplot(summarized_data, aes(x = square_feet, y = total_energy_consumption)) +
  geom_point(color = 'blue', alpha = 0.7) +
  labs(title = "Total Energy Consumption vs Building Size",
       x = "Square Feet of Building",
       y = "Total Energy Consumption") +
  theme_minimal() +
  geom_smooth(method = "lm", col = "red", se = FALSE)  # Adds a linear trend line
```
## Building Metadata

```{r}
# Load the building metadata file
setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")
building_metadata <- read.csv("building_metadata.csv")

# Load the energy consumption file (train or test data)
setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")
energy_data <- read.csv("train 1-2.csv")  # Adjust file path as necessary

# Merge the building metadata and energy data on the 'building_id' column
merged_data <- merge(building_metadata, energy_data, by = "building_id")

# Summarize the total energy consumption by building primary use
summarized_data <- merged_data %>%
  group_by(primary_use) %>%
  summarize(total_energy_consumption = sum(meter_reading, na.rm = TRUE))

# Create a bar plot using ggplot2
ggplot(summarized_data, aes(x = reorder(primary_use, -total_energy_consumption), y = total_energy_consumption)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Total Energy Consumption by Building Type",
       x = "Building Type (Primary Use)",
       y = "Total Energy Consumption") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability
```
## Weather Test

```{r}
setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")

# Load the weather data
weather_data <- read.csv("weather_test.csv")  # Adjust the file path as necessary

# Summarize the average air temperature for each site
summarized_data <- weather_data %>%
  group_by(site_id) %>%
  summarize(average_air_temperature = mean(air_temperature, na.rm = TRUE))

# Create a box plot for air temperature across different sites
ggplot(weather_data, aes(x = factor(site_id), y = air_temperature)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(title = "Distribution of Air Temperature by Site",
       x = "Site ID",
       y = "Air Temperature (°C)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability
```

```{r}
# Scatter plot showing Wind Speed vs Wind Direction
ggplot(weather_data, aes(x = wind_direction, y = wind_speed, color = factor(site_id))) +
  geom_point(alpha = 0.7) +
  labs(title = "Wind Speed vs Wind Direction by Site",
       x = "Wind Direction (degrees)",
       y = "Wind Speed (m/s)") +
  theme_minimal()
```
## Training the Data
```{r}
setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")
install.packages("randomForest")
```


```{r}
install.packages("caret")
install.packages("randomForest")
# Load necessary libraries
library(dplyr)
library(ggplot2)


# Load the datasets
setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")
building_metadata <- read.csv("building_metadata.csv")
train_data <- read.csv("train 1-2.csv")
weather_train <- read.csv("weather_train.csv")

# Merge datasets by building_id and site_id
train_merged <- merge(train_data, building_metadata, by = "building_id")
train_merged <- merge(train_merged, weather_train, by = c("site_id", "timestamp"))

# Convert timestamp to datetime and extract features (hour, day, month)
train_merged$timestamp <- as.POSIXct(train_merged$timestamp)
train_merged$hour <- format(train_merged$timestamp, "%H")
train_merged$day <- format(train_merged$timestamp, "%d")
train_merged$month <- format(train_merged$timestamp, "%m")

# Select relevant features for the model
features <- c("square_feet", "year_built", "floor_count", "air_temperature", 
              "dew_temperature", "precip_depth_1_hr", "cloud_coverage", 
              "sea_level_pressure", "wind_speed", "hour", "day", "month")

# Remove missing data (if necessary)
train_merged <- na.omit(train_merged)
X <- train_merged[, features]
y <- train_merged$meter_reading


```


#### let's start fresh

```{r}
# Load libraries
library(ggplot2)
library(dplyr)

# Load data
setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")
building_metadata <- read.csv("building_metadata.csv")
test_data <- read.csv("test 1-2.csv")
train_data <- read.csv("train 1-2.csv")
weather_test <- read.csv("weather_test.csv")
weather_train <- read.csv("weather_train.csv")

```
```{r}
head(building_metadata)
summary(building_metadata)
str(building_metadata)

```


```{r}
# Load data
setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")
building_metadata <- read.csv("building_metadata.csv")
test_data <- read.csv("test1_2.csv")
train_data <- read.csv("train1_2.csv")
weather_test <- read.csv("weather_test.csv")
weather_train1 <- read.csv("weather_train1.csv")
```
```{r}
# Load necessary libraries
install.packages("scales")
library(ggplot2)
library(dplyr)

# Merge train data with building metadata to link building types
train_with_metadata <- merge(train_data, building_metadata, by = "building_id", all.x = TRUE)

# Check for NA values and drop rows with missing "primary_use" or "meter_reading"
train_with_metadata <- train_with_metadata %>%
  filter(!is.na(primary_use) & !is.na(meter_reading))

# Aggregate meter readings by primary_use and calculate the total energy usage
energy_by_building_type <- train_with_metadata %>%
  group_by(primary_use) %>%
  summarise(total_energy_usage = sum(meter_reading)) %>%
  arrange(desc(total_energy_usage))

# Plotting
ggplot(energy_by_building_type, aes(x = total_energy_usage, y = reorder(primary_use, total_energy_usage))) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "Total Energy Usage by Building Type",
    x = "Total Energy Usage (Meter Reading)",
    y = "Building Type (Primary Use)"
  ) +
  theme_minimal(base_size = 14)


```
```{r}
# Convert total energy usage to kWh
energy_by_building_type <- train_with_metadata %>%
  group_by(primary_use) %>%
  summarise(total_energy_usage = sum(meter_reading) / 1000) %>% # Convert to kWh
  arrange(desc(total_energy_usage))

# Plotting with different colors
ggplot(energy_by_building_type, aes(x = total_energy_usage, y = reorder(primary_use, total_energy_usage), fill = primary_use)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Total Energy Usage by Building Type",
    x = "Total Energy Usage (kWh)", # Add units to the x-axis
    y = "Building Type (Primary Use)"
  ) +
  theme_minimal(base_size = 14) +
  scale_x_continuous(labels = scales::comma) + # Format numbers with commas
  theme(legend.position = "none") # Optionally hide legend if it's not needed
```


```{r}
library(scales) # For label formatting

# Filter data to include valid square_feet, year_built, and meter_reading
valid_data <- train_with_metadata %>%
  filter(!is.na(square_feet) & !is.na(year_built) & !is.na(meter_reading))

# Create a scatter plot to visualize the relationship
ggplot(valid_data, aes(x = square_feet, y = meter_reading, color = year_built)) +
  geom_point(alpha = 0.7, size = 1) + # Set point transparency and size
  scale_color_viridis_c() + # Use the Viridis color scale
  scale_x_continuous(
    breaks = seq(0, 800000, by = 200000), # Set x-axis breaks
    labels = label_comma() # Format labels with commas
  ) +
  labs(
    title = "Square Feet, Year Built, vs. Energy Usage",
    x = "Building Size (Square Feet)", 
    y = "Energy Usage (Meter Reading)",
    color = "Year Built"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")
```
```{r}
# Load data
setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")
building_metadata <- read.csv("building_metadata.csv")
test1_2 <- read.csv("test1_2.csv")
train_data <- read.csv("train1_2.csv")
weather_test <- read.csv("weather_test.csv")
weather_train1 <- read.csv("weather_train1.csv")
```



```{r}
test1_2 <- data.frame(
  meter = factor(rep(1:3, each = 10)),
  timestamp = seq.POSIXt(from = as.POSIXct("2023-01-01"), by = "day", length.out = 30),
  meter_reading = runif(30, 50, 500)
)
```

```{r}
test1_2$meter <- as.factor(test1_2$meter)

ggplot(test1_2, aes(x = timestamp, y = meter_reading, color = meter, group = meter)) +
  geom_line() +
  labs(
    title = "Daily Energy Usage Trends by Meter Type",
    x = "Date",
    y = "Meter Reading (Aggregate)",
    color = "Meter Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  ) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_datetime(date_labels = "%Y-%m-%d")

```
```{r}
# Ensure the 'meter' column includes all levels (0, 1, 2, 3)
test1_2$meter <- factor(test1_2$meter, levels = c("0", "1", "2", "3"))

# Plotting the data
ggplot(test1_2, aes(x = timestamp, y = meter_reading, color = meter, group = meter)) +
  geom_line() +
  labs(
    title = "Daily Energy Usage Trends by Meter Type",
    x = "Date",
    y = "Meter Reading (Aggregate)",
    color = "Meter Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  ) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_datetime(date_labels = "%Y-%m-%d")

```
```{r}
summary(test1_2$timestamp)
head(test1_2$timestamp)

test1_2$timestamp <- as.POSIXct(test1_2$timestamp, format = "%Y-%m-%d %H:%M:%S")
test1_2$timestamp <- as.POSIXct(test1_2$timestamp, format = "%Y-%m-%d")
sum(is.na(test1_2$timestamp))
test1_2 <- test1_2[!is.na(test1_2$timestamp), ]

# Generate proper timestamps for each meter type
test1_2 <- data.frame(
  meter = factor(rep(c("0", "1", "2", "3"), each = 10)),
  timestamp = rep(seq.POSIXt(from = as.POSIXct("2023-01-01"), by = "day", length.out = 10), 4),
  meter_reading = runif(40, 50, 500)
)

ggplot(test1_2, aes(x = timestamp, y = meter_reading, color = meter, group = meter)) +
  geom_line() +
  labs(
    title = "Daily Energy Usage Trends by Meter Type",
    x = "Date",
    y = "Meter Reading (Aggregate)",
    color = "Meter Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  ) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_datetime(date_labels = "%Y-%m-%d")

```
```{r}
# Correct the year to 2016 in the timestamp column
test1_2$timestamp <- as.POSIXct(test1_2$timestamp, format = "%Y-%m-%d %H:%M:%S")
test1_2$timestamp <- as.POSIXct(format(test1_2$timestamp, "2016-%m-%d %H:%M:%S"))

```

```{r}
# Generate proper timestamps for each meter type with year 2016
test1_2 <- data.frame(
  meter = factor(rep(c("0", "1", "2", "3"), each = 10)),
  timestamp = rep(seq.POSIXt(from = as.POSIXct("2016-01-01"), by = "day", length.out = 10), 4),
  meter_reading = runif(40, 50, 500)
)

```

```{r}
library(ggplot2)
library(scales)

ggplot(test1_2, aes(x = timestamp, y = meter_reading, color = meter, group = meter)) +
  geom_line() +
  labs(
    title = "Daily Energy Usage Trends by Meter Type (2016)",
    x = "Date",
    y = "Meter Reading (Aggregate)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  ) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_datetime(date_labels = "%Y-%m-%d") +
  facet_wrap(~ meter, ncol = 2)  # Create a 2x2 grid of plots

```

```{r}
library(ggplot2)
library(scales)

# List of unique meter types
meters <- unique(test1_2$meter)

# Loop through each meter type and create individual plots
for (meter_type in meters) {
  # Filter data for the current meter type
  meter_data <- subset(test1_2, meter == meter_type)
  
  # Create the plot
  p <- ggplot(meter_data, aes(x = timestamp, y = meter_reading)) +
    geom_line(color = "blue") +
    labs(
      title = paste("Daily Energy Usage Trend for Meter", meter_type, "(2016)"),
      x = "Date",
      y = "Meter Reading (Aggregate)"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    scale_y_continuous(labels = scales::comma) +
    scale_x_datetime(date_labels = "%Y-%m-%d")
  
  # Print the plot
  print(p)
}

```

```{r}
library(ggplot2)
library(scales)

# List of unique meter types
meters <- unique(test1_2$meter)

# Assign colors for each meter type
meter_colors <- c("0" = "red", "1" = "blue", "2" = "green", "3" = "purple")

# Loop through each meter type and create individual plots
for (meter_type in meters) {
  # Filter data for the current meter type
  meter_data <- subset(test1_2, meter == meter_type)
  
  # Create the plot
  p <- ggplot(meter_data, aes(x = timestamp, y = meter_reading)) +
    geom_line(color = meter_colors[as.character(meter_type)]) +
    labs(
      title = paste("Daily Energy Usage Trend for Meter", meter_type, "(2016)"),
      x = "Date",
      y = "Meter Reading (Aggregate)"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    scale_y_continuous(labels = scales::comma) +
    scale_x_datetime(date_labels = "%Y-%m-%d")
  
  # Print the plot
  print(p)
}

```

```{r}
# Install tidyr if not already installed
if (!requireNamespace("tidyr", quietly = TRUE)) {
    install.packages("tidyr")
}

# Load the required library
library(tidyr)

```

```{r}
options(repos = c(CRAN = "https://cran.rstudio.com/"))
install.packages("naniar")
install.packages("path_to_downloaded_package.tar.gz", repos = NULL, type = "source")

```
```{r}
library(ggplot2)

# Convert missing data to logical matrix
missing_data <- as.data.frame(is.na(merged_data))

# Create a heatmap using ggplot2
ggplot(melt(missing_data), aes(Var2, Var1, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "white"), name = "Missing") +
  labs(title = "Missing Data Heatmap", x = "Columns", y = "Rows") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r}
install.packages("naniar")

# Install and load required packages
if (!requireNamespace("naniar", quietly = TRUE)) {
  install.packages("naniar")
}
```

```{r}
install.packages("naniar")
library(naniar)
library(ggplot2)

# Visualize missing data
gg_miss_var(merged_data) +
  labs(title = "Missing Data Heatmap by Variable", x = "Variables", y = "Number of Missing Values")

# Alternatively, a full heatmap of missing data
ggplot(data = as.data.frame(is.na(merged_data)), aes(x = 1:ncol(merged_data), y = 1:nrow(merged_data))) +
  geom_tile(aes(fill = as.logical(unlist(merged_data)))) +
  scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "white")) +
  labs(title = "Heatmap of Missing Data", x = "Columns", y = "Rows") +
  theme_minimal()

```{r}
merged_data <- merged_data %>% drop_na(primary_use, square_feet, year_built, floor_count, meter_reading)

```

```{r}
# Inspect the structure of the data
str(merged_data)

# Check for missing values in all relevant columns
summary(merged_data)

# Specifically check the `year_built` column for non-numeric or missing values
table(is.na(merged_data$year_built))



```
```{r}
# Remove rows with missing or invalid values in `year_built` or other relevant columns
merged_data <- merged_data %>%
  filter(!is.na(year_built), !is.na(primary_use), !is.na(square_feet), !is.na(floor_count), !is.na(meter_reading))

# Ensure `year_built` is numeric
merged_data$year_built <- as.numeric(merged_data$year_built)

```
```{r}
# Group year_built into bins for better visualization
merged_data <- merged_data %>%
  mutate(year_built_group = cut(year_built, breaks = 10, include.lowest = TRUE))

# Plot Year Built vs Meter Reading
ggplot(merged_data, aes(x = year_built_group, y = meter_reading)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(title = "Year Built vs. Meter Reading", x = "Year Built (Binned)", y = "Meter Reading (Log Scale)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r}
# Inspect cleaned data
summary(cleaned_data)
head(cleaned_data)

# Check for problematic values
table(is.na(cleaned_data$square_feet))
table(is.na(cleaned_data$meter_reading))

```


```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Set working directory and load data
setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")

# Load datasets
building_metadata <- read.csv("building_metadata.csv")
test1_2 <- read.csv("test1_2.csv")
train_data <- read.csv("train1_2.csv")
weather_test <- read.csv("weather_test.csv")
weather_train1_2 <- read.csv("weather_train1_2.csv")

# Merge building metadata with train data
data <- merge(building_metadata, train_data, by = "building_id", all.x = TRUE)

# Clean and filter the data
data_clean <- data %>%
  filter(!is.na(square_feet), !is.na(meter_reading)) %>%  # Remove missing values
  filter(square_feet > 0, meter_reading > 0)             # Remove zero or negative values

# Plot Square Feet vs Meter Reading
ggplot(data_clean, aes(x = square_feet, y = meter_reading)) +
  geom_point(alpha = 0.5, color = "orange") +            # Add scatter points with transparency
  scale_x_log10() +                                      # Apply log scale to x-axis
  scale_y_log10() +                                      # Apply log scale to y-axis
  labs(
    title = "Square Feet vs. Meter Reading",
    x = "Square Feet (Log Scale)",
    y = "Meter Reading (Log Scale)"
  ) +
  theme_minimal()                                        # Use a clean, minimal theme

```
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(scales) # For formatting axis labels

# Ensure data is clean (replace 'data' with your actual dataframe name)
data_clean <- data %>%
  filter(!is.na(square_feet), !is.na(meter_reading)) %>%  # Remove missing values
  filter(square_feet > 0, meter_reading > 0)             # Remove zero or negative values

# Plot Square Feet vs Meter Reading with formatted axis labels
ggplot(data_clean, aes(x = square_feet, y = meter_reading)) +
  geom_point(alpha = 0.5, color = "orange") +            # Add scatter points
  scale_x_log10(labels = comma) +                        # Log scale for x-axis with formatted labels
  scale_y_log10(labels = comma) +                        # Log scale for y-axis with formatted labels
  labs(
    title = "Square Feet vs. Meter Reading",
    x = "Square Feet (Log Scale)",
    y = "Meter Reading (Log Scale)"
  ) +
  theme_minimal()                                        # Minimal theme for a clean look

```

```{r}
# Load datasets


ggplot(data_clean, aes(x = meter_reading)) +
    geom_histogram(binwidth = 10, fill = "blue", color = "white", alpha = 0.7) +
    scale_x_log10() +
    labs(title = "Distribution of Meter Reading", x = "Meter Reading (Log Scale)", y = "Frequency")

```
```{r}
ggplot(data_clean, aes(x = as.factor(meter), y = meter_reading)) +
    geom_boxplot() +
    scale_y_log10() +
    labs(title = "Meter Type vs. Meter Reading", x = "Meter Type", y = "Meter Reading (Log Scale)")

```
```{r}
ggplot(data_clean, aes(x = primary_use, y = meter_reading)) +
    geom_boxplot() +
    scale_y_log10() +
    labs(title = "Primary Use vs. Meter Reading", x = "Primary Use", y = "Meter Reading (Log Scale)") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
ggplot(data_clean, aes(x = cut(year_built, breaks = 10), y = meter_reading)) +
    geom_boxplot() +
    scale_y_log10() +
    labs(title = "Year Built vs. Meter Reading", x = "Year Built (Binned)", y = "Meter Reading (Log Scale)") + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
ggplot(data_clean, aes(x = as.Date(timestamp), y = meter_reading, group = 1)) +
    geom_line() +
    scale_y_log10() +
    labs(title = "Energy Consumption Over Time", x = "Time", y = "Meter Reading (Log Scale)")

```
```{r}
ggplot(data_clean, aes(x = as.factor(floor_count), y = meter_reading)) +
    geom_boxplot() +
    scale_y_log10() +
    labs(title = "Floor Count vs. Meter Reading", x = "Floor Count", y = "Meter Reading (Log Scale)")

```
```{r}
library(reshape2)
library(corrplot)

# Select numerical columns
numeric_data <- data_clean %>% select(square_feet, meter_reading, year_built, floor_count)
correlation_matrix <- cor(numeric_data, use = "complete.obs")

# Plot heatmap
corrplot(correlation_matrix, method = "color", tl.col = "black", tl.cex = 0.8, addCoef.col = "black")

```
```{r}
library(lubridate)

# Convert timestamp to Date
data_clean$timestamp <- as.Date(data_clean$timestamp)

# Calculate daily averages
daily_avg <- data_clean %>%
    group_by(timestamp) %>%
    summarise(avg_meter_reading = mean(meter_reading, na.rm = TRUE))

# Plot daily average energy consumption
ggplot(daily_avg, aes(x = timestamp, y = avg_meter_reading)) +
    geom_line(color = "blue") +
    scale_y_log10() +
    labs(title = "Daily Average Energy Consumption", x = "Date", y = "Average Meter Reading (Log Scale)") +
    theme_minimal()

```
```{r}
ggplot(data_clean, aes(x = as.Date(timestamp), y = meter_reading, color = as.factor(meter))) +
    geom_line(alpha = 0.7) +
    scale_y_log10() +
    facet_wrap(~ meter) +
    labs(title = "Energy Consumption by Meter Type Over Time", x = "Date", y = "Meter Reading (Log Scale)", color = "Meter Type") +
    theme_minimal()

```

```{r}
ggplot(data_clean, aes(x = as.factor(site_id), y = meter_reading)) +
    geom_boxplot(fill = "purple", alpha = 0.7) +
    scale_y_log10() +
    labs(title = "Energy Consumption Across Sites", x = "Site ID", y = "Meter Reading (Log Scale)") +
    theme_minimal()



```

```{r}
# Load necessary libraries
library(ggplot2)
library(reshape2)

# Assuming 'correlation_matrix' is a data frame or matrix containing the correlations
# Convert the correlation matrix to a format suitable for ggplot
correlation_data <- melt(correlation_matrix)

# Plot the heatmap
ggplot(data = correlation_data, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, 
                       limits = c(-1, 1), space = "Lab", name = "Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation Matrix Heatmap", x = "Variables", y = "Variables")

```

```{r}
# Load necessary libraries
library(dplyr)

setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")
building_metadata <- read.csv("building_metadata.csv")
train_data <- read.csv("train1_2.csv")
weather_train <- read.csv("weather_train1_2.csv")

# Assuming the datasets are already loaded:
# building_metadata <- read.csv("building_metadata.csv")
# train_data <- read.csv("train1_2.csv")
# weather_train <- read.csv("weather_train1_2.csv")

# Merge building metadata with training data
train_merged <- train_data %>%
  left_join(building_metadata, by = "building_id")

# Convert timestamps to datetime format for merging with weather data
train_merged$timestamp <- as.POSIXct(train_merged$timestamp, format = "%m/%d/%Y %H:%M")
weather_train$timestamp <- as.POSIXct(weather_train$timestamp, format = "%m/%d/%Y %H:%M")

# Merge weather data with training data based on site_id and timestamp
train_merged <- train_merged %>%
  left_join(weather_train, by = c("site_id", "timestamp"))

# Add derived columns
train_merged <- train_merged %>%
  mutate(
    energy_per_sqft = meter_reading / square_feet,
    building_age = 2024 - year_built
  )

# Group and aggregate data
agg_data <- train_merged %>%
  group_by(primary_use, building_age, floor_count) %>%
  summarise(
    avg_energy_usage = mean(meter_reading, na.rm = TRUE),
    avg_energy_per_sqft = mean(energy_per_sqft, na.rm = TRUE),
    avg_temp = mean(air_temperature, na.rm = TRUE)
  )

# Print the aggregated data
print(agg_data)
```
```{r}
library(ggplot2)

setwd("C:/Users/Khalil/Desktop/Fall 24/CPT_S 575/Project/Data")


# Load the dataset (update the path as necessary)
data <- read.csv("Aggregated_Energy_and_Building_Characteristics_Data.csv")

# Sort the data by average energy usage per square foot
data_sorted <- data[order(-data$avg_energy_per_sqft), ]

# Create the bar plot using ggplot2
ggplot(data_sorted, aes(x = reorder(primary_use, -avg_energy_per_sqft), y = avg_energy_per_sqft)) +
  geom_bar(stat = "identity", fill = "dark blue") +
  theme_minimal() +
  labs(
    title = "Energy Usage per Square Foot by Building Type",
    x = "Building Type (Primary Use)",
    y = "Average Energy Usage per Square Foot"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


